Are Lower Income Communities Given More Noise Exemptions

### 2024-09-24 paper/paper.qmd
fixed confound section

@@

```{r}
#| label: tbl-wardexemptionssummary
#| tab-cap: "Noise Exemptions Summary And Outliers"
#| fig-subcap: ["Summary", "Outliers"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

Based on the @fig-wardtotalexemptionsscatterplots. There is not seem to be a negative relation ship between median ward income and number of exemptions. However, by looking at the plot, there does seem to me a moderate postive relationship between median income and number of exemptions for all types of permit, although the constrution permit plot's relationship does apear to be weaker.

It is important to note that if we hypothetically removed the outliers (ward 10, 11, 13), the relationship would appear to be a bit weaker, but the relationship still seems reasonable.

@@

```{r}
#| label: fig-confoudingplots
#| fig-cap: "Ward Median Income Relationship With Common Potential Confounds"
#| fig-subcap: ["TODO", "TODO"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

correlation_data <- noise_exemp_ward_data %>% 
  select(ward, ward_area_km2, ward_population, ward_median_income)

ward_income_pop <- correlation_data %>% 
  ggplot(aes(ward_population, ward_median_income, )) +
  geom_point(alpha = 0.7) +
  theme_linedraw() +
  labs(y = "Median Income", x = "Population") + 
  geom_text(aes(label = ward), vjust = 1.5, hjust = 0.5, size = 2)

ward_income_area <- correlation_data %>% 
  ggplot(aes(ward_area_km2, ward_median_income)) +
  geom_point(alpha = 0.7) +
  theme_linedraw() +
  labs(y = "Median Income", x = "ward area (kms squared)") + 
  geom_text(aes(label = ward), vjust = 1.5, hjust = 0.5, size = 2)

ward_income_pop
ward_income_area
```
@fig-confoudingplots explores if median income can be predicted by a ward's population size or area. Based on the plots, there does not appear to be a reltionship between ward median income and the ward's area. As for population, there does seem to be a moderate to weak relationship between ward population and median income.

Nevertheless, if we hypotectically remove the outliers from population graph. Then there is no relationship between population and median income.

>>>>

TODO: add name and captions

@@

```{r}
#| label: tbl-wardexemptionssummary
#| tab-cap: "Noise Exemptions Summary And Outliers"
#| tbl-subcap: ["Summary", "Outliers"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

Based on the @fig-wardtotalexemptionsscatterplots. There is not seem to be a negative relation ship between median ward income and number of exemptions. However, by looking at the plot, there does seem to me a moderate positive relationship between median income and number of exemptions for all types of permit, although the construction permit plot's relationship does appear to be weaker.

It is important to note that if we hypothetically removed the outliers (ward 10, 11, 13), the relationship would appear to be a bit weaker, but there still appears to be a moderate positive relationship.

@@

```{r}
#| label: fig-confoudingplots
#| fig-cap: "Ward Population impact on Medium Impact"
#| fig-subcap: ["TODO", "TODO"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

pop_exemptions = correlation_data %>% ggplot(aes(ward_population, num_exemptions)) +
  geom_point(alpha = 0.7) +
  theme_linedraw() +
  labs(y = "Exemptions", x = "Population") + 
  geom_text(aes(label = ward), vjust = 1.5, hjust = 0.5, size = 2)

correlation_data_confound <- noise_exemp_ward_data %>% 
  select(ward, ward_area_km2, ward_population, ward_median_income)

ward_income_pop <- correlation_data_confound %>% 
  ggplot(aes(ward_population, ward_median_income)) +
  geom_point(alpha = 0.7) +
  theme_linedraw() +
  labs(y = "Median Income", x = "Population") + 
  geom_text(aes(label = ward), vjust = 1.5, hjust = 0.5, size = 2)

# ward_income_area <- correlation_data %>% 
#   ggplot(aes(ward_area_km2, ward_median_income)) +
#   geom_point(alpha = 0.7) +
#   theme_linedraw() +
#   labs(y = "Median Income", x = "ward area (kms squared)") + 
#   geom_text(aes(label = ward), vjust = 1.5, hjust = 0.5, size = 2)

ward_income_pop
pop_exemptions
# ward_income_area

```
@fig-confoudingplots explores a ward's population is a good predictor of a ward's population and the number of permit exemptions given. Based on the plots there does seem to be a moderate positive relationship between a ward's population and a ward's medium income. Likewise, there is a moderate positive relationship between ward population and the number of noise exemptions a ward has given although this reltionship is not as strong as population and median income.

However, we need to also consider that if we hypothetically remove the outlines from population graph. Then there appeared to be very little if not no relationship between population and median income.
